MR-PFU: A suite of R packages for energy conversion chain analysis

### 2023-07-28 paper.Rmd
Better example
More about length

Max length: 250--1000 words

@@

The most important design decision for the suite of `R` packages
involved data format.
We authors are among those who
developed the Physical Supply-Use Table (PSUT) framework,
a matrix-based approach to describing energy flows 
("products" in PSUT terminology) 
from resource extraction to
processing stages
("industries" in PSUT terminology)
and, ultimately, 
to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018; @Aramendia:2022tv].
We chose the PSUT framework as the data format for the `R` packages, 
because it succinctly describes an entire ECC
for one country and one year
via a set of six matrices
(the "RUVY" matrices).
See the following table.

@@

Broadly speaking, 
four calculation steps are required to create a PFU database. 
Each step is completed by one or more new `R` packages.
First, the IEA's primary- and final-stage WEEB data must be converted
to RUVY matrices for each country and year, 
a task accomplished by functions in the [IEATools][link-ieatools] package [@Heun-IEATools:2023]. 
Second, human and animal muscle work must be calculated from 
[International Labor Organization][link-ilo] (ILO) and
[Food and Agriculture Organization][link-fao] (FAO) data,
following the methodology of @Steenwyk:2022ww, 
using functions in the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
Third, the IEA's primary- and final-stage ECC data 
must be extended to the useful stage by
(a) allocating final stage energy to end-use machines and
(b) multiplying allocated final energy by the 
final-to-useful efficiency of each machine.
This task is accomplished by functions in the
[Recca][link-recca] [@Heun-Recca:2023]
and 
[PFUDatabase][link-pfudatabase] [@Heun-PFUDatabase:2023]
packages.
Fourth, ECCs must be converted from energy terms to exergy terms. 
This step is performed by functions in the [Recca][link-recca] package.

@@

Create a MR-PFU database from the GHA and ZAF data contained in the IEATools package.
Create the expectd folder structure that we use on the database.
Store all necessary data in this repository.
Execute the commands to produce the database, 
albeit for only two countries and two years.

>>>>

Max length: 250--1000 words. At the moment, we are at 1444 words.
But I saw many JOSS papers are longer than 1000 words. 
And our paper is describing many packages, whereas
most JOSS papers describe only a single package.
So I think we are going to be OK.
I think we can own the issue in the cover letter, 
saying that we are willing to consider splitting into
two papers upon the editor's advice.

@@

The most important design decision for the suite of `R` packages
involved data format.
We authors are among those who
developed the Physical Supply-Use Table (PSUT) framework,
a matrix-based approach to describing energy flows 
("products" in PSUT terminology) 
from resource extraction to
processing stages
("industries" in PSUT terminology)
and, ultimately, 
to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018; @Aramendia:2022tv].
We chose the PSUT framework as the data format for the new `R` packages, 
because it succinctly describes an entire ECC
for one country and one year
via a set of six matrices
(the "RUVY" matrices).
See the following table.

@@

Broadly speaking, 
four calculation steps are required to create a MR-PFU database. 
Each step is assisted by functions in one or more new `R` packages.
First, the IEA's primary- and final-stage WEEB data must be converted
to RUVY matrices for each country and year, 
a task accomplished by functions in the [IEATools][link-ieatools] package [@Heun-IEATools:2023]. 
Second, human and animal muscle work must be calculated from 
[International Labor Organization][link-ilo] (ILO) and
[Food and Agriculture Organization][link-fao] (FAO) data,
following the methodology of @Steenwyk:2022ww, 
using functions in the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
Third, the IEA's primary- and final-stage ECC data 
must be extended to the useful stage by
(a) allocating final stage energy to end-use machines and
(b) multiplying allocated final energy by the 
final-to-useful efficiency of each machine.
This task is accomplished by functions in the
[Recca][link-recca] [@Heun-Recca:2023]
and 
[PFUDatabase][link-pfudatabase] [@Heun-PFUDatabase:2023]
packages.
Fourth, ECCs must be converted from energy terms to exergy terms. 
This step is performed by functions in the [Recca][link-recca] package.

@@

Create a MR-PFU database from the GHA and ZAF data contained in the IEATools package.   
Create the folder structure for the MR-PFU database.   
Store all necessary data in this repository.   
Execute the commands to produce the database, 
albeit for only two countries and two years.
